Robustness of MinD oscillation in \emph{Escherichia coli} with diverse cell shapes

### 2014-02-05 paper/paper.tex
more updates to methods and pill section of paper Tested everything on node32.

@@

We begin with the naturally occuring pill cell shape.  We peice this
shape together as two hemispherical endcaps attached on either end of
a cylinder.  This shape follows the early simulations of Huang
\emph{et al.} but differs in that we have added the end caps for a
more natural shape, expecting similar results.  We test cells of
radius $.5\micron$ and of lengths $2\micron$, $3\micron$, and
$5\micron$ measured from the tip of one end cap to the top of the
other. In each case, we observe quick establishement (within a few
periods) of extremely regular oscillatory behavoir that lasts
indefinitely.  The period increases with cell length, as shown in
Table~\ref{tab:pill-periods}, in agreement with the results of Huang
\emph{et al.}.

\begin{table}
  \begin{tabular}{|r|c|c|c|l|}
    \hline
    Length($\mu$) & 2.50 & 3.00 & 4.00 & 5.00\\
    \hline
    Period(sec) & sim & 33 & 38 & 48 \\ \cline{2-2}
    \hline
  \end{tabular}
  \caption{Period of oscillation according to length of cell for
    pill-shaped cells.}\label{tab:pill-periods}
\end{table}

This simple model is a good starting point for observing in detail the
dynamic interaction between the different stages of protein that lead
to their qualitatively distinct behavoir. Figure \ref{image-p} shows
the process in a series of frames taken from simulation animations.
Each frame is 2.5 seconds ahead of the last, and each image shows a
the concentration of a given state of protein summed over the
coordinate orthogonal to the page.  The color scale for each protein
state is set according to the maximum density observed.

At $t=0$ in Figs.~\ref{image-p}, the cytoplasmic MinE
is concentrated in the lower portion of the cell, and is diffusing
upward.  The MinD protein is initially concentrated on the membrane in
the upper portion of the cell.  At the boundary between the
cytoplasmic MinE and the membrane-bound MinD, there is a ring of the
membrane-bound MinE:MinD complex, which is moving towards the upper
end of the cell, removing the MinD from the membrane as it goes.  The
MinD is removed from the membrane in its MinD:ADP form, which begins
to accumulate at the upper end of the cell, peaking peaking at around
10 seconds.

>>>>

We begin with the naturally occuring pill cell shape.  We peice this
shape together as two hemispherical endcaps attached on either end of
a cylinder.  This shape follows the early simulations of Huang
\emph{et al.} but differs in that we have added the end caps for a
more natural shape, expecting similar results.  We test cells of
radius $.5\micron$ and of lengths $2\micron$, $3\micron$, and
$5\micron$ measured from the tip of one end cap to the top of the
other. In each case, we observe quick establishement (within a few
periods) of extremely regular oscillatory behavoir that lasts
indefinitely.  The period increases with cell length, as shown in
Table~\ref{tab:pill-periods}, in agreement with the results of Huang
\emph{et al.}

\begin{table}
  \begin{tabular}{|r|c|c|c|l|}
    \hline
    Length($\mu$) & 2.50 & 3.00 & 4.00 & 5.00\\
    \hline
    Period(sec) & sim & 33 & 38 & 48 \\ \cline{2-2}
    \hline
  \end{tabular}
  \caption{Period of oscillation according to length of cell for
    pill-shaped cells.}\label{tab:pill-periods}\fixme{whats the 'sim'
    here?}
\end{table}

This simple cell shape is a good starting point for observing in
detail the dynamic interaction between the different stages of protein
that lead to their qualitatively distinct behavoir. Figure
\ref{image-p} shows the process in a series of frames taken from
simulation animations.  Each frame is 2.5 seconds ahead of the last,
and each image shows a the concentration of a given state of protein
summed over the coordinate orthogonal to the page.  The color scale
for each protein state is set according to the maximum density
observed.

At $t=0$ in Figs.~\ref{image-p}, the cytoplasmic MinE
is concentrated in the lower portion of the cell, and is diffusing
upward.  The MinD protein is initially concentrated on the membrane in
the upper portion of the cell.  At the boundary between the
cytoplasmic MinE and the membrane-bound MinD, there is a ring of the
membrane-bound MinE:MinD complex, which is moving towards the upper
end of the cell, removing the MinD from the membrane as it goes.  The
MinD is removed from the membrane in its MinD:ADP form, which begins
to accumulate at the upper end of the cell, peaking at around
10 seconds.

@@
